---
output:
  md_document:
    variant: gfm
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# faahydro

<!-- badges: start -->
<!-- badges: end -->

faahydro is an R package for fetching, processing, and analyzing hydrologic and 
water quality data. This package was created to support work in South Florida
by Federico & Associates, Inc.

## Installation

This package is currently only available on github. Use the `devtools::install_github()` function
to install it:

``` r
# install.packages("devtools")
devtools::install_github("walkerjeffd/faahydro")
```

Then load the package

```{r load}
library(faahydro)
```

## DBHYDRO Data

### Fetching Data

The `dbhydro_get_*()` functions are used to fetch hydrologic and water quality data from DBHYDRO.
These functions call the respective `get_*()` functions from the `dbhydroR` package,
and then perform some additional cleaning of the dataset (e.g., parsing dates, removing duplicate columns).

```{r dbhydro-get, eval = FALSE}
# Not Run
dbhydro_get_hydro(
  dbkeys = "91599",
  date_min = "2019-10-01",
  date_max = "2019-10-31"
)
dbhydro_get_wq(
  station_ids = "LOX3",
  date_min = "2019-09-01",
  date_max = "2019-10-31",
  test_name = "PHOSPHATE, TOTAL AS P"
)
```

Due to API limits on DBHYDRO, it is sometimes necessary to fetch data in batches using the `dbhydro_batch_get_*()` functions.

```{r dbhydro-batch-get, eval = FALSE}
# Not Run
dbhydro_batch_get_hydro(
  dbkeys = c("91599", "91473", "91663"),
  date_min = "2019-10-01",
  date_max = "2019-10-31",
  batch_size = 2
)
dbhydro_batch_get_wq(
  station_ids = c("LOX3", "LOX4", "LOX5"),
  date_min = "2019-09-01",
  date_max = "2019-10-31",
  test_name = "PHOSPHATE, TOTAL AS P",
  batch_size = 2
)
```

### Cleaning Data

`TODO`

## Database

SQL scripts to create the database can be found in the `inst/sql` folder.

### Set Up

To set up a new database, first create it.

``` sh
createdb faadb
```

Then install the schema for each set of tables.

``` sh
psql -d faadb -f inst/sql/schema-dbhydro.sql
```


